Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopy

### 2024-07-15 paper.tex
edits

Inspired by recent work which demonstrated the presence of HDFG in ultrafast DOVE experiments,\cite{McDonnell2024} we investigate the parameters which result in both SR and DR-SIVE output. \cite{Cho2000, Bonn2024}
We develop simple expressions which relate transition dipoles and hyper-Raman polarizabilties and the overall HDFG output, yielding gross selection rules. 
It is shown that a HDFG active mode must be both IR and hyper-Raman active.
Since any IR active vibration is also hyper-Raman active, \cite{Andrews1978} we demonstrate that HDFG is a coherent analogue of hyper-Raman scattering spectroscopy for IR active vibrations, i.e., HDFG is an optically upconverted infrared spectroscopy.
This is akin to sum and difference frequency generation (SFG, DFG) being coherent Raman analogues for IR active vibrations. \cite{Shen90}

%todo: discussion on ABC terms and what they mean.

We investigate properties of DR-HDFG to understand how coupling between vibrational and vibronic states can be measured.
DR-HDFG is shown to demand one and two-photon absorptions between the ground and excited state.
DR-HDFG is thus the four wave mixing analogue of doubly resonant infrared-visible sum frequency generation. \cite{Shen94}

\section{Steady State Hyper-DFG Spectroscopy}
Singly resonant hyper difference generation spectroscopy (HDFG) spectroscopy has potential to provide deeper insight into single quantum decoherence times and molecular orientation in condensed systems.
Similarities between HDFG spectroscopy, infrared spectroscopy and spontaneous hyper-Raman scattering have been noted previously. \cite{RN352, Bonn2024}
In this section, we make the connections between HDFG and hyper-Raman scattering explicit and investigate the steady state properties of singly resonant HDFG.

@@

To make the connection between HDFG and hyper-Raman scattering, we investigate its gross selection rules.
By propagating density matrix elements in the steady state limit, the HDFG hyperpolarizability is \cite{RN119}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl} =	- \sum_{m, n} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{vn}_{i} \mu^{nm}_{j} \mu^{mg}_{k} \mu^{gv}_{l} }{\Delta_{nv} \Delta_{mv}\Delta_{gv}}  \rho_{gg}
\end{equation}
where: $\mu^{ab}_{j}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl} = \omega_{kl} - \omega_{j} - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the dephasing of $\rho_{kl}$ and $\rho_{gg}$ is the ground state population.
D, the Maker-Terhune degeneracy factor, accounts for permutation symmetry in $\gamma_{ijkl}$.\cite{RN134} 
For a HDFG experiment using two or three distinct input fields, as considered here, D = 6.
We first investigate the selection rules through a Placzek type formalism.
Contracting over the virtual states forms the hyper-Raman hyperpolarizability $\beta$.\cite{Long1970, Willetts1992} 
Assuming $\omega_2$, $\omega_3$ are significantly detuned from any local resonance (\autoref{fig:comparisonwmel}),\cite{Placzek1934, Long1970} \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar}\frac{\beta^{gv}_{ijk} \mu^{gv}_{l}}{\Delta_{gv}} \rho_{gg}
\end{equation}
Taylor expanding the dipole and first hyperpolarizability operators in terms of an n$^{\text{th}}$ normal mode coordinate $Q_n$ about equilibrium: \cite{Long1970}
\begin{subequations}
	\begin{equation}
		\mu = \mu_0 + \frac{\partial \mu}{\partial Q_n} Q_n + \order{{Q_m Q_n}}
	\end{equation}
	\begin{equation}
		\beta = \beta_{0} + \frac{\partial \beta}{\partial Q_n} Q_n + \order{{Q_m Q_n}}
	\end{equation}
\end{subequations}
Substituting into \autoref{sivebeta} gives the HDFG hyperpolarizability to $\order{Q_n}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{48 m_n \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \frac{\partial \beta^{gv}_{ijk}}{\partial Q_n} {\frac{\partial \mu^{gv}_{l}}{\partial Q_n}}  \rho_{gg}
\end{equation}
where $\mel{v}{Q_n}{g} = \sqrt{\frac{\hbar}{2m_n\omega_{vg}}}$ and $m_n$ is the ``mass'' of the $Q_n$ mode.  \cite{RN230}
Since this expression is non-zero in the harmonic oscillator limit, HDFG output is allowed for harmonic transitions. 
Similar to SFG, HDFG does not demand anharmonic ground state potential energy surfaces for spectral output. \cite{Shen94, Cho2000}
This selection rule is generally valid for any HDFG or hyper sum frequency generation (HSFG) process in the $-\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ and $\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ phasematching geometries when $\omega_2$ and $\omega_3$ are sufficiently detuned from any resonance.

While Placzek approximation type arguments are useful to understand the general source of HDFG output, it is insightful to interpret the origin of signal in terms of Albrecht ABC type terms.\cite{Ziegler1988} 
As such, we extend $\beta_{ijk}$ in \autoref{SIVEselection} using the hyper-Raman ABC formalism to understand how vibronic coupling manipulates selection rules for singly and doubly resonant HDFG \autoref{fig:hdfg}. \cite{Ziegler1988, Baranov1990}

@@

To employ the Albrecht formalism, we write the states in terms of a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a(Q)) \otimes \ket{b}$ for electronic states $\{|a(Q))\}$ and vibrational states $\{\ket{b}\}$ (i.e., adiabatic approximation). \cite{BornOppenheimer, Tang1970}
The dependence of the electronic states on the adiabatic parameter $Q$ is suppressed for simplicity.
The states are labeled as shown in \autoref{fig:hdfg}.
For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments ($\vec{\mu}$ is reserved for transitions between states on the ground electronic state). \cite{Ziegler1974}
Following the approach which gave \autoref{sivegamma}, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,e,v'} \frac{
		R_{i}^{gv, ev'} 
		R_{j}^{ev',mn} 
		R_{k}^{mn,g0} 
		R_{l}^{g0,gv} 
	}{\Delta_{g0,gv}
		\Delta_{ev', mn}
		\Delta_{mn, g0}
	}
\end{equation}
where $R_{i}^{ab,cd}$ is the i$^{th}$ element of $\mel{a,b}{\vec{R}}{c,d}$.\cite{Ziegler1988}
Using our definition of vibronic states, we write $\vec{M}^{ab} = (a|\vec{M}|b)$ so that, for example,
$R_{i}^{gv,ev'} = \mel{v}{M_i^{ge}}{v'}$.
Using these ideas rewrites \autoref{drgamma_notaylor} as
\begin{equation}\label{drgamma_notaylor_1}
	\begin{split}
		\gamma_{ijkl} &= -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,e,v'} \frac{
			\mel{v}{M_{i}^{ge}}{v'} 
			\mel{v'}{M_{j}^{em}}{n}
			\mel{n}{M_{k}^{mg} }{v}
			\mel{v}{M_{l}^{gg}}{0}
		}{\Delta_{g0,gv}
			\Delta_{ev', mn}
			\Delta_{mn, g0}	} \\
%	&= 	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
%		\mel{v}{M_{i}^{ge}}{v'} 
%		\mel{v'}{M_{j}^{em}}{n}
%		\mel{n}{M_{k}^{mg} }{v}
%		\mel{v}{\mu_{l}}{0}
%	}{\Delta_{g0,gv}
%		\Delta_{ev', mn}
%		\Delta_{mn, g0}	} \\
	\end{split}
\end{equation}
Chung and Ziegler showed some years ago that expanding $\vec{M}^{ij}$ to $\order{Q}$ as
\begin{equation}
\vec{M}^{ij} = \vec{M}^{ij}_0 + \sum_z \frac{\partial\vec{M}^{ij}}{\partial Q_z} Q_z
\end{equation} 
yields $A, B, C$ coefficients similar to those in the Albrecht formalism of Raman scattering, \cite{Albrecht1961, Ziegler1988} so that
\begin{equation}
		\gamma_{ijkl} \sim \frac{(A_{ijk} + B_{ijk} + C_{ijk})\mel{v}{\mu_{l}}{0}} {\Delta_{g0,gv}}
\end{equation}
The $A$ term depends upon $\order{Q^0}$ transitions (i.e., Condon approximation), the $B$ term depends upon $\order{Q}$ transitions, and the $C$ term depends on $\order{Q^2}$ transitions. 
The C term is suppressed in the following discussion as it depends on one and two photon forbidden transitions. \cite{Ziegler1988}
Note that other reports obtain $A, B, C$ coefficents through a Herzberg-Teller expansion of the electronic states, where $|a) = |a)_0 + \sum_{a \neq b} \frac{(\partial H / \partial Q )_0  Q}{E_a - E_b} |a)_0$.\cite{HerzbergTeller1933, Neddersen1989, Baranov1990}
This approach is not used here as the expansion of the vibronic dipole moment in terms of normal mode coordinates provides similar insight.

By compiling all the terms that scale with $\order{Q^0}$, $A$ is written, assuming contraction over the nonresonant vibrational states $\ket{n}$ as
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,v'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | v' \rangle
		 \langle v' | 0 \rangle 
		 \frac{1}{\Delta_{ev',m} \Delta_{m, g0}}
		 \\
	\end{split}
\end{equation}
and by compiling all terms that scale with $\order{Q}$, $B$ is written as $B = B_1 + B_2$,\cite{Bonang1992} where
\begin{subequations}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,v',z} M^{ge}_{0,i} \langle v | v' \rangle \lbrack
			 \frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k} \mel{v'}{Q_z}{0} 
			+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z} \mel{v'}{Q_z}{0} \rbrack \frac{1}{\Delta_{ev',m} \Delta_{m, g0}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,v',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{v'} 
			\langle v' | 0 \rangle 
			\frac{1}{\Delta_{ev',m} \Delta_{m, g0}}
	\end{split}
	\end{equation}
\end{subequations}
where terms such as $\langle a | b \rangle$ are Franck-Condon factors.

In the case of singly resonant HDFG, $\omega_3$ is also detuned which further contracts the sum over ${v'}$. 
Assuming a normal mode basis, we see for singly resonant HDFG that
\begin{subequations}
\begin{equation}
	A_{ijk} \sim \sum_{v'}\langle v | v' \rangle \langle v' | 0 \rangle = \delta_{v0} = 0
\end{equation}
\begin{equation}
	B_{1_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,z} M^{ge}_{0,i} \lbrack
	\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}
	+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z} \rbrack \frac{\mel{v}{Q_z}{0}}{\Delta_{e,m} \Delta_{m, g0}}
\end{equation}
\begin{equation}
	B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
	M^{mg}_{0,k}  
	\frac{\mel{v}{Q_z}{0}}{\Delta_{e,m} \Delta_{m, g0}}
\end{equation}
\end{subequations}
i.e., A$_{ijk}$ vanishes, but B$_{ijk}$ is non-zero, for singly resonant HDFG. 
The singly resonant HDFG susceptibility can be approximated as
\begin{equation}
	\chi^{(3)}_{IJKL} \sim \langle B_{ijk} \mu_l \rangle
\end{equation}
or, while a vibrational mode must be infrared active for HDFG output, vibronic activity is a necessity for singly resonant HDFG response.
Breakdown of the Condon approximation in singly resonant HDFG supports the relative small intensities of HDFG, reported to be roughly an order of magnitude larger than nonresonant background, in hexane and other organic solvents. \cite{RN351, RN353}

The selection rules for DR-HDFG are different. 
DR-HDFG output is dependent upon the $A,B$ terms prior to contraction over $\ket{v'}$. 
Based upon the presence of the $A$ term in DR-HDFG, it is apparent that DR-HDFG can inform on molecular electronic structure.
DR-HDFG (where a vibrational and vibronic state are resonantly coupled) can provide a tool for measuring vibronic coupling, analogous to fully resonant DFG. \cite{Dick83_1, Shen94}
Many multi-resonant nonlinear spectroscopies have been developed to resolve vibronic coupling in molecular samples. \cite{Carlson1990, Gaynor2017, RN276}
DR-HDFG provides methods to investigate vibronic coupling with only two laser pulses.

@@

@@

\section{Conclusions}%didn't change to HDFG down here because lazy
Coherent vibrational, hyper-Raman coherent four wave mixing spectroscopies are identified and discussed.
Singly resonant SIVE (SR-SIVE) processes are shown to be the coherent hyper-Raman analogue of infrared active vibrations.
Since SR-SIVE is always non-zero for IR active vibrations, the method could be used as an analytical method to optically upconvert infrared transitions with small transition dipoles.
A method for measuring hyper-Raman polarizabilities, $\beta_{ijk}$, through SIVE spectroscopy was identified and discussed.
Extracting the magnitude and sign of $\beta_{ijk}$ using SIVE methods can help assess the quality of theoretical methods which calculate hyper-Raman spectra.

>>>>

Inspired by recent work which demonstrated the presence of HDFG in ultrafast DOVE experiments,\cite{McDonnell2024} we investigate the parameters which result in both SR and DR-SIVE output. \cite{Cho2000, Bonn2024}
This paper is outlined as follows.
In section \ref{steadystate}, we discuss the selection rules behind singly and doubly resonant HDFG in terms of hyper-Raman A,B,C coefficients.
In section \ref{quant}, the quantitative aspects of HDFG is discussed.
In section \ref{mixeddomain}, mixed time-frequency domain HDFG is discussed, and is applied to the simple case of CH$_3$CN in section \ref{Expt}. 
\section{Experimental}\label{experimental}

The ultrafast laser system has been described in detail elsewhere. \cite{RN278}
Briefly, an 80 MHz Ti-Sapphire ultrafast oscillator  (Spectra-Physics Tsunami) created 35 fs seed pulses, amplified by a 1 kHz regnerative amplifier (Spectra-Physics Spitfire Ace) to approximately 0.5 W.
A mask in the regenerative amplifier stretched the input pulses to roughly 1 ps.
The resultant output was used to pump four independent OPAs, two of which are relevant for this study.
OPA1 (TOPAS-800, Light Conversion), referred to as $\omega_1$, was tuned for difference frequency generation (DFG) with an external stage (NDFG ``DF2'', Light Conversion).
OPA2 (OPA-800C, Spectra Physics), referred to as $\omega_2$, was tuned in its signal arrangement. 
OPA3, a homebuilt OPA, is tuned in its signal arrangement and referred to as $\omega_3$.
Grating and crystal positions in OPA1 was externally controlled using \texttt{WinTOPAS} (Light Conversion).
Grating and BBO crystal positions in OPA2 and OPA3 were computer controlled through communication with DC actuators.
The OPAs were independently tuned through the \texttt{Attune} package, controlled by the \texttt{yaqc} and \texttt{Bluesky} packages. \cite{RN414, RN386, SkyeOPA, KyleOPA}
The power spectra and tuning results are found in Figure X. %todo: actually graph these
Movable retroreflectors (Thorlabs) controlled by 50 mm DC actuators (PMC) are used to measure coherence dynamics.
The beams are focused using a BOXCARS geometry.\cite{RN308, Kaufman2024}
To enforce frequency dependent temporal pulse overlap, the maximum points of a third order cross correlation between $\omega_1$ and $\omega_3$ measured in CaF$_2$ were used to generate frequency dependent zero delay values.
The nonlinear output was isolated through chopping schemes and a set of spatial apertures, directed into a monochromator (Horiba Micro-HR) and homodyne detected through a photomultiplier tube (Hamamatsu H7422-20), recorded and converted to digital output using a data acquisition unit (National Instruments).
\textit{in-situ} power corrections are performed by dividing the recorded spectra by OPA output measured by home-built pyroelectric detectors.
All spectra are reported on the intensity scale.

The cyanocobalamin (CNCbl) [DOT Scientific inc., CAS no. 68-19-9] samples were prepared as drop casted thin films.
CNCbl was dissolved in methanol (MeOH) [Fisher Scientific, HPLC Grade, CAS no. 67-56-1] until a concentrated solution was formed (concentration $\sim$ 10 mg CNCbl/ mL MeOH). 
All chemicals were used as obtained without further purification.
The solution was deposited on a microscope coverslip and allowed to dry in a dessicator until a thin layer was formed.

\section{Steady State HDFG Spectroscopy}\label{steadystate}
Singly resonant hyper difference generation spectroscopy (HDFG) spectroscopy has potential to provide deeper insight into single quantum decoherence times and molecular orientation in condensed systems.
Similarities between HDFG spectroscopy, infrared spectroscopy and spontaneous hyper-Raman scattering have been noted previously. \cite{RN352, Bonn2024}
In this section, we investigate the steady state properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.

@@

To make the connection between HDFG and hyper-Raman scattering, we investigate its gross selection rules.
By propagating density matrix elements in the steady state limit, the HDFG hyperpolarizability is \cite{RN119}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl} =	- \sum_{m, n} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{vn}_{i} \mu^{nm}_{j} \mu^{mg}_{k} \mu^{gv}_{l} }{\Delta_{nv} \Delta_{mv}\Delta_{gv}}  \rho_{gg}
\end{equation}
where: $\mu^{ab}_{j}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl} = \omega_{kl} - \omega_{j} - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the dephasing of $\rho_{kl}$ and $\rho_{gg}$ is the ground state population.
D, the Maker-Terhune degeneracy factor, accounts for permutation symmetry in $\gamma_{ijkl}$.\cite{RN134} 
For a HDFG experiment using two or three distinct input fields, as considered here, D = 6.
We first investigate the selection rules through a Placzek type formalism.
Contracting over the virtual states forms the hyper-Raman hyperpolarizability $\beta$.\cite{Long1970, Willetts1992} 
Assuming $\omega_2$, $\omega_3$ are significantly detuned from any local resonance (\autoref{fig:comparisonwmel}),\cite{Placzek1934, Long1970} \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar}\frac{\beta^{gv}_{ijk} \mu^{gv}_{l}}{\Delta_{gv}} \rho_{gg}
\end{equation}
It is important to note that the properties of $\beta_{ijk}$ change dependent upon whether $\omega_2$ and $\omega_3$ are degenerate. \cite{Andrews1978, Altmann1982}
Taylor expanding the dipole and first hyperpolarizability operators in terms of an n$^{\text{th}}$ normal mode coordinate $Q_n$ about equilibrium as\cite{Long1970}
\begin{subequations}
	\begin{equation}
		\mu = \mu_0 + \frac{\partial \mu}{\partial Q_n} Q_n + \order{{Q_m Q_n}}
	\end{equation}
	\begin{equation}
		\beta = \beta_{0} + \frac{\partial \beta}{\partial Q_n} Q_n + \order{{Q_m Q_n}}
	\end{equation}
\end{subequations}
and substituting into \autoref{sivebeta} gives the HDFG hyperpolarizability to $\order{Q_n}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{48 m_n \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \frac{\partial \beta^{gv}_{ijk}}{\partial Q_n} {\frac{\partial \mu^{gv}_{l}}{\partial Q_n}}  \rho_{gg}
\end{equation}
where $\mel{v}{Q_n}{g} = \sqrt{\frac{\hbar}{2m_n\omega_{vg}}}$ and $m_n$ is the ``mass'' of the $Q_n$ mode.  \cite{RN230}
Since this expression is non-zero in the harmonic oscillator limit, HDFG output is allowed for harmonic transitions. 
Similar to SFG, HDFG does not demand anharmonic ground state potential energy surfaces for spectral output. \cite{Shen94, Cho2000}
This selection rule is generally valid for any HDFG or hyper sum frequency generation (HSFG) process in the $-\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ and $\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ phasematching geometries when $\omega_2$ and $\omega_3$ are sufficiently detuned from any resonance.

@@

While Placzek type arguments provide general source of HDFG output, it is insightful to interpret the origin of signal in terms of Albrecht ABC type terms.\cite{Ziegler1988} 
We write $\beta_{ijk}$ in \autoref{SIVEselection} using the hyper-Raman ABC formalism to understand how vibronic coupling manipulates selection rules for singly and doubly resonant HDFG (\autoref{fig:hdfg}). \cite{Ziegler1988, Baranov1990}
To employ this formalism, we write the states in terms of a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a(Q)) \otimes \ket{b}$ for electronic states $\{|a(Q))\}$ and vibrational states $\{\ket{b}\}$ (i.e., adiabatic approximation). \cite{BornOppenheimer, Tang1970}
The dependence of the electronic states on the adiabatic parameter $Q$ is suppressed for simplicity.
The states are labeled as shown in \autoref{fig:hdfg}.
For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments ($\vec{\mu}$ is reserved for transitions between states on the ground electronic state). \cite{Tang1970}
Following the approach which gave \autoref{sivegamma}, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,e,v'} \frac{
		R_{i}^{gv, ev'} 
		R_{j}^{ev',mn} 
		R_{k}^{mn,g0} 
		R_{l}^{g0,gv} 
	}{\Delta_{g0,gv}
		\Delta_{ev', mn}
		\Delta_{mn, g0}
	}
\end{equation}
where $R_{i}^{ab,cd}$ is the i$^{th}$ element of $\mel{a,b}{\vec{R}}{c,d}$.
Using our definition of vibronic states, we write $\vec{M}^{ab} = (a|\vec{M}|b)$ so that, for example,
$R_{i}^{gv,ev'} = \mel{v}{M_i^{ge}}{v'}$.
Using these ideas rewrites \autoref{drgamma_notaylor} as
\begin{equation}\label{drgamma_notaylor_1}
	\begin{split}
		\gamma_{ijkl} &= -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,e,v'} \frac{
			\mel{v}{M_{i}^{ge}}{v'} 
			\mel{v'}{M_{j}^{em}}{n}
			\mel{n}{M_{k}^{mg} }{v}
			\mel{v}{\mu_{l}}{0}
		}{\Delta_{g0,gv}
			\Delta_{ev', mn}
			\Delta_{mn, g0}	} \\
%	&= 	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
%		\mel{v}{M_{i}^{ge}}{v'} 
%		\mel{v'}{M_{j}^{em}}{n}
%		\mel{n}{M_{k}^{mg} }{v}
%		\mel{v}{\mu_{l}}{0}
%	}{\Delta_{g0,gv}
%		\Delta_{ev', mn}
%		\Delta_{mn, g0}	} \\
	\end{split}
\end{equation}
Chung and Ziegler showed some years ago that expanding $\vec{M}^{ij}$ to $\order{Q}$ as
\begin{equation}
\vec{M}^{ij} = \vec{M}^{ij}_0 + \sum_z \frac{\partial\vec{M}^{ij}}{\partial Q_z} Q_z
\end{equation} 
yields $A, B, C$ coefficients similar to those in the Albrecht formalism of Raman scattering, \cite{Albrecht1961, Ziegler1988} so that
\begin{equation}
		\gamma_{ijkl} \sim \frac{(A_{ijk} + B_{ijk} + C_{ijk})\mel{v}{\mu_{l}}{0}} {\Delta_{g0,gv}}
\end{equation}
The $A$ term depends upon $\order{Q^0}$ transitions (i.e., Condon approximation), the $B$ term depends upon $\order{Q}$ transitions, and the $C$ term depends on $\order{Q^2}$ transitions. 
The C term is suppressed in the following discussion as it depends on one and two photon forbidden transitions. \cite{Ziegler1988, Neddersen1989, Bonang1992}
Note that some reports obtain $A, B, C$ coefficients through a Herzberg-Teller expansion of the electronic states, $|a) = |a)_0 + \sum_{a \neq b} \frac{(\partial H / \partial Q )_0  Q}{E_a - E_b} |a)_0$, to explicitly expose vibronic couplings in terms of $\partial H / \partial Q $, where H is the electronic Hamiltonian.\cite{HerzbergTeller1933, Petrov1985, Neddersen1989, Baranov1990}
This approach is not used here as the expansion of $\vec{M}^{ij}$ in normal mode coordinates provides sufficient physical insight into hyper-Raman selection rules.

By contracting over the virtual vibrational states $\ket{n}$, the $A$ and $B$ coefficients, where $B = B_1 + B_2$, are written as
\begin{widetext}
\begin{subequations}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,v'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | v' \rangle
		 \langle v' | 0 \rangle 
		 \frac{1}{\Delta_{ev',m} \Delta_{m, g0}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,v',z} M^{ge}_{0,i} \langle v | v' \rangle \lbrack
			 \frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k} \mel{v'}{Q_z}{0} 
			+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z} \mel{v'}{Q_z}{0} \rbrack \frac{1}{\Delta_{ev',m} \Delta_{m, g0}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,v',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{v'} 
			\langle v' | 0 \rangle 
			\frac{1}{\Delta_{ev',m} \Delta_{m, g0}}
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
where terms such as $\langle a | b \rangle$ are Franck-Condon factors. 
These expressions are valid for both singly and doubly resonant HDFG, but based upon the resonance schemes, further simplifications can be made (\textit{vide infra}).

In the case of singly resonant HDFG, $\omega_3$ is also detuned from any resonance, which permits contraction over ${v'}$ as the dependence on $\ket{v'}$ in the resonance denominator can be neglected. 
Assuming a normal mode basis, we see for singly resonant HDFG that
\begin{subequations}
\begin{equation}
	A_{ijk} \sim \sum_{v'}\langle v | v' \rangle \langle v' | 0 \rangle = \delta_{v0} = 0
\end{equation}
\begin{equation}
	B_{1_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,z} M^{ge}_{0,i} \lbrack
	\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}
	+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z} \rbrack \frac{\mel{v}{Q_z}{0}}{\Delta_{e,m} \Delta_{m, g0}}
\end{equation}
\begin{equation}
	B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
	M^{mg}_{0,k}  
	\frac{\mel{v}{Q_z}{0}}{\Delta_{e,m} \Delta_{m, g0}}
\end{equation}
\end{subequations}
i.e., A$_{ijk}$ vanishes, but B$_{ijk}$ is non-zero, for singly resonant HDFG. 
The singly resonant HDFG susceptibility can be approximated as
\begin{equation}
	\chi^{(3)}_{IJKL} \sim \langle B_{ijk} \mu_l \rangle
\end{equation}
or, while a vibrational mode must be infrared active for HDFG output, vibronic activity is necessary for singly resonant HDFG response.
The breakdown of the Condon approximation in singly resonant HDFG supports the relative small HDFG $\chi^{(3)}$ values ($\sim$ order of magnitude larger than nonresonant background) in hexane and similar organic solvents. \cite{RN350, RN351, RN353}

The selection rules for DR-HDFG are different. 
DR-HDFG (where a vibrational and vibronic state are resonantly coupled) can provide a tool for measuring vibronic coupling, analogous to fully resonant DFG. \cite{Dick83_1, Shen94}
Many multi-resonant nonlinear spectroscopies have been developed to resolve vibronic coupling in molecular samples. \cite{Carlson1990, Gaynor2017, RN276}
DR-HDFG provides methods to investigate vibronic coupling with only two laser pulses.
DR-HDFG output is dependent upon the $A,B$ terms prior to contraction over $\ket{v'}$. 
Based upon the presence of the $A$ term in DR-HDFG, it is apparent that DR-HDFG can inform on molecular electronic structure.

@@

\section{Quantitative Aspects of HDFG}\label{quant}

@@

\section{Conclusions}%todo:fix!
Coherent vibrational, hyper-Raman coherent four wave mixing spectroscopies are identified and discussed.
Singly resonant SIVE (SR-SIVE) processes are shown to be the coherent hyper-Raman analogue of infrared active vibrations.
Since SR-SIVE is always non-zero for IR active vibrations, the method could be used as an analytical method to optically upconvert infrared transitions with small transition dipoles.
A method for measuring hyper-Raman polarizabilities, $\beta_{ijk}$, through SIVE spectroscopy was identified and discussed.
Extracting the magnitude and sign of $\beta_{ijk}$ using SIVE methods can help assess the quality of theoretical methods which calculate hyper-Raman spectra.
